My title

### 2024-03-13 paper/paper.qmd
updates on results section, removed the model r script as it was causing error somehow.

```{r}
#| echo: false
#| message: false
#| warning: false
model <- survey_model

# Assuming 'model' is your model object from rstanarm
library(broom.mixed)

coefficients <- broom.mixed::tidy(model, conf.int = TRUE)

# Assuming you have knitr and kableExtra already installed and loaded
library(knitr)
library(kableExtra)

knitr::kable(coefficients,
             caption = "Coefficients from the Model",
             digits = 2) %>%
  kableExtra::kable_styling(latex_options = "scale_down")

@@

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

@@

# get proportions by demographic, then find proportion of each cell relative to population
post_props <- post_strat_data %>% 
  group_by(state, race, age_group, gender, education) %>% 
  summarise(n = n(), .groups ='drop') %>% 
  group_by(state) %>% 
  mutate(prop = n/sum(n))

# generate predictions for each of our grouped cells
post_props$estimate <- predict.glm(survey_model_now, newdata = post_props,
                          type = "response")

# create a new column with number of voters from each cell voting Trump
post_props <- post_props %>% mutate(num_voters = n*estimate)

# get errors for our prediction
errors <- predict.glm(survey_model_now, newdata = post_props,
                      type = "response", se.fit = T)
lower <- errors$fit - errors$se.fit
upper <- errors$fit + errors$se.fit

# combine the proportions and the errors we generated
post_props_error <- cbind(post_props, lower, upper)
post_props_error <- post_props_error %>% 
  rename("lower" = ...10,
         "upper" = ...11)

# create a new column with number of voters from each cell voting Biden
post_props_error <- post_props_error %>% mutate(num_voters_lower = n*lower,
                                num_voters_upper = n*upper)

# create a new dataset which contains upper and lower estimates for proportions
# of votes for each state
vote_trump <- post_props_error %>% mutate(trump_prop=estimate*prop,
                                       trump_prop_lower = lower*prop,
                                       trump_prop_upper = upper*prop) %>%
  group_by(state) %>%
  summarise(trump_predict = sum(trump_prop),
            trump_prop_lower = sum(trump_prop_lower),
            trump_prop_upper = sum(trump_prop_upper))

@@

```{r stateplot} 
#| echo: false 
#| message: false
#| label: fig-modelresults
#| fig-cap: "Proportion of each State Voting for Trump"
#| warning: false

# get state by state proportions for our polling data
survey_props <- survey_data %>% 
  group_by(state, consider_trump_binary) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))

# combine the polling data proportions with the post-stratification data
combined_trump <- inner_join(survey_props, vote_trump, by = "state")
combined_trump  <- combined_trump %>% 
  filter(consider_trump_binary == 1)

# arrange our proportions for each state in descending order

@@

# plot the results, with a line on 50% because of the electoral college

stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
                               color = support)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
  labs(title = "Estimated Proportion of State Voting for Joe Biden",
       x = "Percentage", y= "State",
       color = "Support") +
  scale_x_continuous(labels = scales::percent) +
  theme(axis.text.y = element_text(size = 5)) +
  scale_color_manual(values = c("Blue", "Red"))
```

Figure \@ref(fig:stateplot) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms. Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state. Horizontal Lines extending from the dots represent confidence intervals for these estimates. The length of each line indicates the uncertainty associated with each estimate. For instance, we can see that this uncertainty lies between 50 percent to slightly higher than 80 percent for Trump in MT (Massachusetts). The dashed green line in the middle at the 50 percent mark represents the threshold for majority support. On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest at the top to the lowest at the bottom.

From figure \@ref(fig:stateplot), it seems that majority of the states support Biden. Only 7 states out of 51 have its point estimate greater than 50 percent for Trump. However, the horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans. However, excluding these contesting states, our model suggests that only 3 states are definitely in favor of Trump whereas 
35 states are definitely supporting Biden.

```{r educationplot} 
#| echo: false 
#| message: false 
#| warning: false 
#| fig-cap: "How different education levels of voters affect voting for Trump"

@@

Figure \@ref(fig:educationplot) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'. Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump. The horizontal lines extending to the left and right of each dot represent the confidence intervals around the estimate, which reflect the uncertainty.

>>>>

@@

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

```{r}
#| echo: false
#| message: false
#| warning: false

@@

# group by each demographic
post_props <- post_strat_data %>% 
  group_by(state, race, age_group, gender, education) %>% 
  summarise(n = n(), .groups ='drop') %>% 
  group_by(state) %>% 
  mutate(prop = n/sum(n))

# estimate of the model 
post_props$estimate <- predict.glm(survey_model_now, newdata = post_props,
                          type = "response")

# new column with the number of voters for Trump
post_props <- post_props %>% mutate(num_voters = n*estimate)

# error terms 
errors <- predict.glm(survey_model_now, newdata = post_props,
                      type = "response", se.fit = T)
lower <- errors$fit - errors$se.fit
upper <- errors$fit + errors$se.fit

# combine errors with proportions
post_props_error <- cbind(post_props, lower, upper)
post_props_error <- post_props_error %>% 
  rename("lower" = ...10,
         "upper" = ...11)

# new column of voters for trump 
post_props_error <- post_props_error %>% mutate(num_voters_lower = n*lower,
                                num_voters_upper = n*upper)

# confidence intervals for errors 
vote_trump <- post_props_error %>% mutate(trump_prop=estimate*prop,
                                       trump_prop_lower = lower*prop,
                                       trump_prop_upper = upper*prop) %>%
  group_by(state) %>%
  summarise(trump_predict = sum(trump_prop),
            trump_prop_lower = sum(trump_prop_lower),
            trump_prop_upper = sum(trump_prop_upper))

@@

```{r} 
#| echo: false
#| message: false
#| warning: false
#| label: fig-distribution-by-state
#| fig-cap: Distribution of Sample and Population by Education

# state proportions for survey data
survey_props <- survey_data %>% 
  group_by(state, consider_trump_binary) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))

# combining survey data with the post-stratification data
combined_trump <- inner_join(survey_props, vote_trump, by = "state")
combined_trump  <- combined_trump %>% 
  filter(consider_trump_binary == 1)

@@

# make a plot of the result for state 
stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
                               color = support)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
  labs(title = "Estimated Proportion of State Voting for Joe Biden",
       x = "Percentage", y= "State",
       color = "Support") +
  scale_x_continuous(labels = scales::percent) +
  theme(axis.text.y = element_text(size = 5)) +
  scale_color_manual(values = c("Blue", "Red"))
```

Figure \@ref(fig:fig-distribution-by-state) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms. Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state. Horizontal Lines extending from the dots represent confidence intervals for these estimates. The length of each line indicates the uncertainty associated with each estimate. For instance, we can see that this uncertainty lies between 50 percent to slightly higher than 80 percent for Trump in MT (Massachusetts). The dashed green line in the middle at the 50 percent mark represents the threshold for majority support. On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest at the top to the lowest at the bottom.

From figure, it seems that majority of the states support Biden. Only 7 states out of 51 have its point estimate greater than 50 percent for Trump. However, the horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans. However, excluding these contesting states, our model suggests that only 3 states are definitely in favor of Trump whereas 
35 states are definitely supporting Biden.

```{r} 
#| echo: false
#| message: false
#| warning: false
#| label: fig-distribution-by-educationlevel
#| fig-cap: How different education levels of voters affect voting for Trump

@@

Figure \@ref(fig:fig-distribution-by-educationlevel) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'. Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump. The horizontal lines extending to the left and right of each dot represent the confidence intervals around the estimate, which reflect the uncertainty.
